Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic

### 2021-06-29 write-up/covid-writeup.tex
Update covid-writeup.tex

\begin{abstract}
  Coronavirus case-count data has influenced government policies and drives most epidemiological forecasts. Limited testing is cited as the key driver behind minimal information on the COVID-19 pandemic. While expanded testing is laudable, measurement error and selection bias are the two greatest problems limiting our understanding of the COVID-19 pandemic; neither can be fully addressed by increased testing capacity. In this paper, we demonstrate their impact on estimation of point prevalence and the effective reproduction number. We show that estimates based on the millions of molecular tests in the US has the same mean square error as a small simple random sample.  To address this, a procedure is presented that combines case-count data and random samples over time to estimate selection propensities based on key covariate information. We then combine these selection propensities with epidemiological forecast models using death-only data to construct a \emph{doubly-robust} estimation method that accounts for both measurement-error and selection bias.  This method is then applied to estimate Indiana's prevalence using case-count, hospitalization, and death data with cumulative demographic information, the United States's only statewide random molecular sample collected from April 25--29th, and Facebook's COVID-19 symptom survey.  We end with a series of recommendations based on the proposed methodology.
\end{abstract}

@@

Selection bias in case count data is addressed by bringing in a different source of information, COVID-19 death data.  Let $D_{k,r}$ denote the number of individuals who pass away on day $r$ from strata $k$.  Then
$$
D_{k,r} \mid p, \theta, \nu \sim \text{Poisson} \left( \sum_{t=1}^r p_{k} \nu_{k,t} \theta_{(r-t)} \right)
$$
where $p_{k}$ is the infection fatality rate for the $k$th strata.  This extends prior analysis of death data~\cite{Johndrow2020} by allowing the infection fatality rate to depend on age.  Indeed,~\cite{Levin2020} demonstrate that ``the estimated IFR is close to zero for children and younger adults but rises exponentially with age, reaching 0.4\% at age 55, 1.4\% at age 65, 4.6\% at age 75, and 15\% at age 85.'' For conciseness, parameter choices and inference are discussed in depth in Appendix~\ref{app:modelbased}.

@@

\begin{lemma}[Variance of doubly-robust estimator] \normalfont
\label{lemma:dr}
\textcolor{red}{Must fix.}
The estimates~$\mu_t := \bar y_{n,t}^\star$ and $\hat \pi_{j,t}$
are solutions to the following set of estimating equations
$$
\Phi_n (\eta_t) =
\left (
\begin{array}{c}
\frac{1}{N} \sum_{j=1}^N I_{j,t} \frac{Y_{j,t} - (1-FP-FN) \cdot (\mu_t+FP)}{\pi_{j,t}} \\
\frac{1}{N} \sum_{t^\prime = 1}^T K_{t,t^\prime} \sum_{j=1}^N I_{j,t^\prime} X_{j,t^\prime} - \frac{1}{N} \sum_{j = 1}^N \tilde I_{j,t^\prime} \tilde W_{j,t^\prime}  \pi_{j,t^\prime} X_{j,t^\prime}  \\
\frac{1}{n_{FP}} \sum_{j \in S_{FP}} \frac{Z_j}{FP} - \frac{1-Z_j}{1-FP} \\
\frac{1}{n_{FN}} \sum_{j \in S_{FN}} \frac{\tilde Z_j}{FN} - \frac{1-\tilde Z_j}{1-FN} \\
\end{array}
\right ) = {\bf 0}
$$
where $\eta_t = (\mu_t, \pi_t, FP, FN)$. Under regularity assumptions (see Appendix~\ref{app:asympderivations}), we have $\bar y_{n,t}^\star - \bar Y_{t} = O_p (n^{-1/2})$ (check)  and $\text{var} (\hat y_{n,t}) = V_{t}^{(IPW)} + o (n^{-1})$ where $V_t^{(IPW)}$ is the first diagonal element of $E [\phi_n(\eta_0)]^{-1} \text{Var}(\Phi_n(\eta_0))E [\phi_n(\eta_0)]^{-1}$.
\end{lemma}

%\section{COVID-19 active infection prevalence in Indiana}
%\label{section:applications}
%% For COVID-19 case-count data, the only covariate information provided by state and local governments are summary statistics for those factors discussed in Section~\ref{section:testinginfo} -- age, sex,  and race.  These are discrete factors and therefore define a stratification of the population.  Currently, US testing does not collect additional covariate information such as symptom information; however, this need not be the case moving forward.  Recommendation XX below states that additional covariate information should be collected from nonprobabilistic samples.
%Here we consider estimation of the active infection rate in Indiana. We consider five sources of data:
%\begin{itemize}[leftmargin=*]
%\item {\bf Testing data}:  Each day, the number of tests performed and number of positive tests are reported, with counts broken out jointly by age, gender, and racial demographic information.  Figure~\ref{fig:in-cases} plots the number of COVID-19 positive cases based on testing date by age.
%\begin{figure}
%\centering
%\begin{subfigure}{.45\textwidth}
%  \centering
%  \includegraphics[width=.9\linewidth]{../figs/indianacasecounts.png}
%  \caption{Indiana case-counts per day}
%  \label{fig:in-cases}
%\end{subfigure}
%\begin{subfigure}{.45\textwidth}
%  \centering
%  \includegraphics[width=.9\linewidth]{../figs/indianadeaths.png}
%  \caption{Indiana death-counts per day}
%  \label{fig:in-deaths}
%\end{subfigure}
%\end{figure}
%\item {\bf Random/Nonrandom statewide sample}: Table~\ref{tab:indiana} summarizes data from a random sample from April 25--29th as well as a nonrandom sample obtained between May 2--3 in racial/ethnic minority communities (cite). The nonrandom sample is not a random subsample of the overall case-counts; however, this data provides important supplementary covariate information.  Namely, the samples collected included covariate information on (1) prior symptoms, (2) COVID-19 positive individual in household, and (3) prior COVID-19 positive test. To account for the nonrandom sample being from high risk areas, we adjust estimates to match the statewide positivity rate of 11.7\% between May 1--2. These adjusted rates are presented in parentheses in Table~\ref{tab:indiana}.
%\begin{table}[th]
%\begin{tabular}{c | c | r r | r | r r }
%& & \multicolumn{2}{c}{Total Tests (\%)} & US & \multicolumn{2}{c}{Positive Test Rate (\%)}\\
%\cline{3-4} \cline{6-7}
%& & NonRandom & Random & Census & NonRandom & Random \\ \hline
%\multirow{2}{*}{Sex} & Female & 58.2 & 55.0 & 50.7 & 21.7 (11.2) & 1.4 \\
%& Male & 41.8 & 45.0 & 49.3 & 24.2 (12.4) & 2.1 \\ \hline
%\multirow{3}{*}{Age} & $<40$ & 39.4 & 28.0 & 52.7 & 29.7 (15.0) & 1.7 \\
%& $40-59$ & 41.1 & 36.0 & 25.2 & 24.9 (12.5) & 2.1 \\
%& $\geq 60$ & 19.5 & 36.0 & 22 & 6.7 (3.4) & 0.9 \\ \hline
%\multirow{2}{*}{Race} & White & 23.1 & 92.0 & 86.9 & 19.5 (9.6) & 1.5 \\
%& Nonwhite & 76.9 & 8.0 & 13.1 & 25.0 (12.3) & 3.4 \\ \hline
%% Hispanic & Hispanic & 44.1 & 2.0 & 7.1 & 37.6 & 6.9 \\
%% Origin & Non-Hispanic & 55.9 & 98.0 & 92.9 & 13.0 & 1.3 \\ \hline
%\multirow{2}{*}{Fever} & Yes & 17.0 & 1.8 & - & 66.4 (32.1) & 4.5 \\
%& No & 83.0 & 98.2 & - & 15.6 (7.5) & 1.3 \\ \hline
%Household & Yes & 10.8 & 1.4 & - & 46.1 & 29.4 \\
%$+$ Case & No & 89.2 & 98.6 & - & 21.6 & 1.3 \\ \hline
%Prior $+$ & Yes & 6.1 & 1.4 & - & 39.2 & 24.4 \\
%Test & No & 93.9 & 98.6 & - & 21.6 & 1.3 \\ \hline
%\end{tabular}
%\caption{Estimated point prevalence of active infection with SARS-CoV-2 by demographics in Indiana.}
%\label{tab:indiana}
%\end{table}
%\item {\bf Facebook symptom survey}: We extract daily symptom surveys completed by individuals on Facebook who identify as living in Indiana.  We collect age, sex, and demographic information as well as COVID-19 related symptoms, which Facebook defines as having (1) a fever and cough, (2) shortness of breath, \emph{or} (3) difficulty breathing. To assess potential bias, we compare the random sample from Indiana to the Facebook symptom survey data collected from April 20th to April 25th.  Table~\ref{tab:comparison} demonstrates minimal bias in the Facebook symptom survey.
%\begin{table}
%\centering
%\begin{tabular}{c | c c c | c}
%& \multicolumn{3}{c}{Facebook} & Random \\ \cline{2-5}
%& Lower CI & Estimate & Upper CI & Estimate \\ \hline
%Fever &  0.017 & 0.020 & 0.024 & 0.018 \\
%Cough &  0.142 & 0.148 & 0.154 & 0.149 \\
%Shortness & 0.058 & 0.060 & 0.062 & 0.062 \\ \hline
%\end{tabular}
%\caption{Comparison of Facebook symptom survey and random sample demonstrates minimal bias.}
%\label{tab:comparison}
%\end{table}
%% \item {\bf Hospitalization data}: On each day, we observe the number of individuals who are COVID-19 positive that visited a hospital on that day.  Unfortunately, hospitalization data is not broken out by age, gender, and racial demographic information.  Instead, we rely on combining Facebook data to calculate the fraction of those who sought the hospital per week in each strata. We then use this to figure out how many individuals per strata were hospitalized and then back out how many
%\item {\bf Death data}: On each day, we observe the number of deaths due to COVID-19 reported on that day, with deaths broken out jointly by age, gender, and racial demographic information.  Figure~\ref{fig:in-deaths} plots the cumulative number of COVID-19 related deaths over time by age group.
%\end{itemize}
%\subsection{Inverse-probability weighting approach}
%We start with using Table~\ref{tab:indiana} to compute IPW weights.  The random sample is $n=3,658$.  Here, we demonstrate how to compute the weight for a particular strata.  For example, the strata defined as Male, 40--59, White, has a Fever, no household cases, and no prior tests has a weight proportional to
%$$
%\frac{\left( 0.493 \times 0.252 \times 0.869 \times 0.018 \right)}{\left(0.418 \times 0.411 \times 0.231 \times 0.17 \right)} = 0.628
%$$
%Using the constructed weights and the adjusted strata-level prevalence estimates, the IPW estimate under no measurement-error is $7.7$, a decrease of four percentage points from the observed prevalence of $11.7$. Using Table~\ref{tab:indiana} to compute weights requires the assumption of independence among factors.  To alleviate this, we compute weights using Facebook's survey data collected in Indiana over that window of time; the resulting IPW estimate under no measurement-error is
%The study did not report sensitivity and specificity; therefore, we take the suggested measures from~\cite{Woloshin2020} which report 70\% sensitivity is a reasonable estimates and~\cite{Cohen2020} which report 97.6\% specificity.  This corresponds to a false negative rate of $30\%$ and false positive rate of $2.4\%$.  Therefore, the IPW estimate accounting for measurement-error is $7.7\%$.  To account for uncertainty in
%Indiana's population as of 2019 was $6.732$ million.  Subtracting off the $617,555$ tests that had already performed yields a sampling fraction of $f = 0.001$.  The smoothed prevalence estimate is 28.4\%.  Adjusting for imperfect testing yields an estimate of $33.9\%$.
%For this estimate, we consider estimation of data quality $E_{\I} \left[ \rho_{\tilde I(X),Y} \right]$. In \cite{Meng2018}, estimation relied on observing the true outcome (i.e., election vote totals), which is not possible in the current crisis.  Here, we use the random sample as a noisy but unbiased estimate of the true prevalence. Our goal is not inferential. Instead, we aim to build sensitivity analyses that can aid in understanding the amount of information in observational COVID case-count data.
%Under the assumption that the survey is a SRS from New York, the error is $17.6$\%.  Using~\eqref{eq:error}, we construct an estimate of the relative sampling rate as follows
%\[
%\rho D_M = \sqrt{\frac{f}{1-f}} \frac{\text{0.176}}{\sigma_Y} = 1.07 \times 10^{-2} \Rightarrow \Delta = 1.07 \times 10^{-3} \Rightarrow M = 2.3.
%\]
%We can perform a sensitivity analysis by considering the range of false negative and positive rates from~\cite{Bendavid2020}, which leads to a range for $M$ of $(2.21, 2.38)$.  Therefore, the impact on bias of $\hat R_t$ appears moderate.  If $FP = 0.05$ and $FN = 0.005$, then the relative sampling rate increases to $M=3.04$.  This simple tool helps us understand how sensitivity and specificity impact biases and  therefore how much we should trust conclusions based on these assumptions.

@@

>>>>

\begin{abstract}
  Coronavirus case-count data has influenced government policies and drives most epidemiological forecasts. Limited testing is cited as the key driver behind minimal information on the COVID-19 pandemic. While expanded testing is laudable, measurement error and selection bias are the two greatest problems limiting our understanding of the COVID-19 pandemic; neither can be fully addressed by increased testing capacity. In this paper, we demonstrate their impact on estimation of point prevalence and the effective reproduction number. We show that estimates based on the millions of molecular tests in the US has the same mean square error as a small simple random sample.  To address this, a procedure is presented that combines case-count data and random samples over time to estimate selection propensities based on key covariate information. We then combine these selection propensities with epidemiological forecast models using death-only data to construct a \emph{doubly robust} estimation method that accounts for both measurement-error and selection bias.  This method is then applied to estimate Indiana's prevalence using case-count, hospitalization, and death data with cumulative demographic information, the United States's only statewide random molecular sample collected from April 25--29th, and Facebook's COVID-19 symptom survey.  We end with a series of recommendations based on the proposed methodology.
\end{abstract}

@@

Selection bias in case count data is addressed by bringing in a different source of information, COVID-19 death data.  Let $D_{k,r}$ denote the number of individuals who pass away on day $r$ from strata $k$.  Then
$$
D_{k,r} \mid p, \theta, \nu \sim \text{Poisson} \left( \sum_{t=1}^r p_{k} \nu_{k,t} \theta_{(r-t)} \right)
$$
where $p_{k}$ is the infection fatality rate for the $k$th strata.  This extends prior analysis of death data~\cite{Johndrow2020} by allowing the infection fatality rate to depend on age.  Indeed,~\cite{Levin2020} demonstrate that ``the estimated IFR is close to zero for children and younger adults but rises exponentially with age, reaching 0.4\% at age 55, 1.4\% at age 65, 4.6\% at age 75, and 15\% at age 85.''  See Appendix~\ref{app:ifr} for details on choice of $p_k$.

% 0-17: 20/1000000 * 100 -> 0.002 (at age 9)
% 18-49: 500/1000000 * 100 -> 0.05 (at age 33)
% 50–64 years old: 6,000/1000000 * 100 -> 0.6
% 65+ years old: 90,000/1000000 * 100 -> 9.0

% For conciseness, parameter choices and inference are discussed in depth in Appendix~\ref{app:modelbased}.

@@

\begin{lemma}[Variance of doubly-robust estimator] \normalfont
\label{lemma:dr}
The estimates~$\mu_t := \bar y_{n,t}^\star$ and $\hat \pi_{j,t}$
are solutions to the following set of estimating equations
$$
\Phi_n (\eta_t) =
\left (
\begin{array}{c}
\frac{1}{N} \left[ \sum_{j=1}^N \frac{I_{j,t}}{\pi_{j,t}} \left( \left( \frac{Y_{j,t} - FP}{1-FP-FN} - \mu(x_j)  \right)  - \mu_t \right) + \frac{1}{\sum_{j=1}^N \tilde I_j \tilde W_j} \sum_{j=1}^N  \tilde I_j \tilde W_j \hat \mu (x_j) \right] \\
\frac{1}{N} \sum_{t^\prime = 1}^T K_{t,t^\prime} \sum_{j=1}^N I_{j,t^\prime} X_{j,t^\prime} - \frac{1}{N} \sum_{j = 1}^N \tilde I_{j,t^\prime} \tilde W_{j,t^\prime}  \pi_{j,t^\prime} X_{j,t^\prime}  \\
\frac{1}{n_{FP}} \sum_{j \in S_{FP}} \frac{Z_j}{FP} - \frac{1-Z_j}{1-FP} \\
\frac{1}{n_{FN}} \sum_{j \in S_{FN}} \frac{\tilde Z_j}{FN} - \frac{1-\tilde Z_j}{1-FN} \\
\end{array}
\right ) = {\bf 0}
$$
where $\eta_t = (\mu_t, \pi_t, FP, FN)$. Under regularity assumptions (see Appendix~\ref{app:asympderivations}), we have $\bar y_{n,t}^\star - \bar Y_{t} = O_p (n^{-1/2})$ (check)  and $\text{var} (\hat y_{n,t}) = V_{t}^{(IPW)} + o (n^{-1})$ where $V_t^{(IPW)}$ is the first diagonal element of $E [\phi_n(\eta_0)]^{-1} \text{Var}(\Phi_n(\eta_0))E [\phi_n(\eta_0)]^{-1}$.
\end{lemma}

\section{COVID-19 active infection prevalence in Indiana}
\label{section:applications}

% For COVID-19 case-count data, the only covariate information provided by state and local governments are summary statistics for those factors discussed in Section~\ref{section:testinginfo} -- age, sex,  and race.  These are discrete factors and therefore define a stratification of the population.  Currently, US testing does not collect additional covariate information such as symptom information; however, this need not be the case moving forward.  Recommendation XX below states that additional covariate information should be collected from nonprobabilistic samples.

Here we consider estimation of the active infection rate in Indiana. We consider five sources of data:
\begin{itemize}[leftmargin=*]
\item {\bf Testing data}:  Each day, the number of tests performed and number of positive tests are reported, with counts broken out jointly by age, gender, and racial demographic information.  Figure~\ref{fig:in-cases} plots the number of COVID-19 positive cases based on testing date by age.

\begin{figure}
\centering
\begin{subfigure}{.45\textwidth}
 \centering
 \includegraphics[width=.9\linewidth]{../figs/indianacasecounts.png}
 \caption{Indiana case-counts per day}
 \label{fig:in-cases}
\end{subfigure}
\begin{subfigure}{.45\textwidth}
 \centering
 \includegraphics[width=.9\linewidth]{../figs/indianadeaths.png}
 \caption{Indiana death-counts per day}
 \label{fig:in-deaths}
\end{subfigure}
\end{figure}

\item {\bf Random/Nonrandom statewide sample}: Table~\ref{tab:indiana} summarizes data from a random sample from April 25--29th as well as a nonrandom sample obtained between May 2--3 in racial/ethnic minority communities~\cite{Yiannoutsos2021}. The nonrandom sample is not a random subsample of the overall case-counts; however, this data provides important supplementary covariate information.  Namely, the samples collected included covariate information on (1) prior symptoms, (2) COVID-19 positive individual in household, and (3) prior COVID-19 positive test. To account for the nonrandom sample being from high risk areas, we adjust estimates to match the statewide positivity rate of 11.7\% between May 1--2. These adjusted rates are presented in parentheses in Table~\ref{tab:indiana}.

\begin{table}[th]
\begin{tabular}{c | c | r r | r | r r }
& & \multicolumn{2}{c}{Total Tests (\%)} & US & \multicolumn{2}{c}{Positive Test Rate (\%)}\\
\cline{3-4} \cline{6-7}
& & NonRandom & Random & Census & NonRandom & Random \\ \hline
\multirow{2}{*}{Sex} & Female & 58.2 & 55.0 & 50.7 & 21.7 (11.2) & 1.4 \\
& Male & 41.8 & 45.0 & 49.3 & 24.2 (12.4) & 2.1 \\ \hline
\multirow{3}{*}{Age} & $<40$ & 39.4 & 28.0 & 52.7 & 29.7 (15.0) & 1.7 \\
& $40-59$ & 41.1 & 36.0 & 25.2 & 24.9 (12.5) & 2.1 \\
& $\geq 60$ & 19.5 & 36.0 & 22 & 6.7 (3.4) & 0.9 \\ \hline
\multirow{2}{*}{Race} & White & 23.1 & 92.0 & 86.9 & 19.5 (9.6) & 1.5 \\
& Nonwhite & 76.9 & 8.0 & 13.1 & 25.0 (12.3) & 3.4 \\ \hline
% Hispanic & Hispanic & 44.1 & 2.0 & 7.1 & 37.6 & 6.9 \\
% Origin & Non-Hispanic & 55.9 & 98.0 & 92.9 & 13.0 & 1.3 \\ \hline
\multirow{2}{*}{Fever} & Yes & 17.0 & 1.8 & - & 66.4 (32.1) & 4.5 \\
& No & 83.0 & 98.2 & - & 15.6 (7.5) & 1.3 \\ \hline
Household & Yes & 10.8 & 1.4 & - & 46.1 & 29.4 \\
$+$ Case & No & 89.2 & 98.6 & - & 21.6 & 1.3 \\ \hline
Prior $+$ & Yes & 6.1 & 1.4 & - & 39.2 & 24.4 \\
Test & No & 93.9 & 98.6 & - & 21.6 & 1.3 \\ \hline
\end{tabular}
\caption{Estimated point prevalence of active infection with SARS-CoV-2 by demographics in Indiana.}
\label{tab:indiana}
\end{table}

\item {\bf Facebook symptom survey}: We extract daily symptom surveys completed by individuals on Facebook who identify as living in Indiana.  We collect age, sex, and demographic information as well as COVID-19 related symptoms, which Facebook defines as having (1) a fever and cough, (2) shortness of breath, \emph{or} (3) difficulty breathing. To assess potential bias, we compare the random sample from Indiana to the Facebook symptom survey data collected from April 20th to April 25th.  Table~\ref{tab:comparison} demonstrates minimal bias in the Facebook symptom survey.
\begin{table}
\centering
\begin{tabular}{c | c c c | c}
& \multicolumn{3}{c}{Facebook} & Random \\ \cline{2-5}
& Lower CI & Estimate & Upper CI & Estimate \\ \hline
Fever &  0.017 & 0.020 & 0.024 & 0.018 \\
Cough &  0.142 & 0.148 & 0.154 & 0.149 \\
Shortness & 0.058 & 0.060 & 0.062 & 0.062 \\ \hline
\end{tabular}
\caption{Comparison of Facebook symptom survey and random sample demonstrates minimal bias.}
\label{tab:comparison}
\end{table}

% \item {\bf Hospitalization data}: On each day, we observe the number of individuals who are COVID-19 positive that visited a hospital on that day.  Unfortunately, hospitalization data is not broken out by age, gender, and racial demographic information.  Instead, we rely on combining Facebook data to calculate the fraction of those who sought the hospital per week in each strata. We then use this to figure out how many individuals per strata were hospitalized and then back out how many

\item {\bf Death data}: On each day, we observe the number of deaths due to COVID-19 reported on that day, with deaths broken out jointly by age, gender, and racial demographic information.  Figure~\ref{fig:in-deaths} plots the cumulative number of COVID-19 related deaths over time by age group.
\end{itemize}

\subsection{Inverse-probability weighting approach}

We start with using Table~\ref{tab:indiana} to compute IPW weights.  The random sample is $n=3,658$.  Here, we demonstrate how to compute the weight for a particular strata.  For example, the strata defined as Male, 40--59, White, has a Fever, no household cases, and no prior tests has a weight proportional to
$$
\frac{\left( 0.493 \times 0.252 \times 0.869 \times 0.018 \right)}{\left(0.418 \times 0.411 \times 0.231 \times 0.17 \right)} = 0.628
$$
Using the constructed weights and the adjusted strata-level prevalence estimates, the IPW estimate under no measurement-error is $7.7$, a decrease of four percentage points from the observed prevalence of $11.7$. Using Table~\ref{tab:indiana} to compute weights requires the assumption of independence among factors.  To alleviate this, we compute weights using Facebook's survey data collected in Indiana over that window of time; the resulting IPW estimate under no measurement-error is

The study did not report sensitivity and specificity; therefore, we take the suggested measures from~\cite{Woloshin2020} which report 70\% sensitivity is a reasonable estimates and~\cite{Cohen2020} which report 97.6\% specificity.  This corresponds to a false negative rate of $30\%$ and false positive rate of $2.4\%$.  Therefore, the IPW estimate accounting for measurement-error is $7.7\%$.  To account for uncertainty in

Indiana's population as of 2019 was $6.732$ million.  Subtracting off the $617,555$ tests that had already performed yields a sampling fraction of $f = 0.001$.  The smoothed prevalence estimate is 28.4\%.  Adjusting for imperfect testing yields an estimate of $33.9\%$.

For this estimate, we consider estimation of data quality $E_{\I} \left[ \rho_{\tilde I(X),Y} \right]$. In \cite{Meng2018}, estimation relied on observing the true outcome (i.e., election vote totals), which is not possible in the current crisis.  Here, we use the random sample as a noisy but unbiased estimate of the true prevalence. Our goal is not inferential. Instead, we aim to build sensitivity analyses that can aid in understanding the amount of information in observational COVID case-count data.

Under the assumption that the survey is a SRS from New York, the error is $17.6$\%.  Using~\eqref{eq:error}, we construct an estimate of the relative sampling rate as follows
\[
\rho D_M = \sqrt{\frac{f}{1-f}} \frac{\text{0.176}}{\sigma_Y} = 1.07 \times 10^{-2} \Rightarrow \Delta = 1.07 \times 10^{-3} \Rightarrow M = 2.3.
\]
We can perform a sensitivity analysis by considering the range of false negative and positive rates from~\cite{Bendavid2020}, which leads to a range for $M$ of $(2.21, 2.38)$.  Therefore, the impact on bias of $\hat R_t$ appears moderate.  If $FP = 0.05$ and $FN = 0.005$, then the relative sampling rate increases to $M=3.04$.  This simple tool helps us understand how sensitivity and specificity impact biases and  therefore how much we should trust conclusions based on these assumptions.

@@

\section{Infection fatality rate (IFR) by age group}
\label{app:ifr}
